A pandemic-scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-09-26 latex/main.tex
Updates from Overleaf

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
We reconstructed a more comprehensive mutational landscape
of SARS-CoV-2 by mapping deletions onto the ARG
using post-hoc parsimony.
We identified important deletion events
(which are inherited by at least 10,000 samples;
Table~\ref{tab:major_dels}).
All these deletions are known to occur major VOCs
(https://virological.org/t/preliminary-genomic-characterisation-of-an-emergent-sars-cov-2-lineage-in-the-uk-defined-by-a-novel-set-of-spike-mutations/563;
\cite{Akaishi2023,Li2023indels}).
Eight of these deletions occur in S1-NTD,
a known hotspot for recurrent deletions \cite{McCarthy2021}.
There are three contiguous deletions in NSP6
(4-base at 11288, 5-base at 11283, and 5-base at 11292),
which should have been mapped as two 9-base deletions,
one above the BA.1 origin node and
the other above the BA.2 origin node in the ARG
(Figure~\ref{fig:subgraph_ba1_ba2}),
but they were broken up
during deletion remapping because they had
overlapping four sites at 11288-11292
(which was mapped as the 4-base deletion event).
Furthermore, the placement of these deletion events in the ARG
reflects that of the VOC-associated deletion events
inferred in a previous study \cite{Akaishi2023}.

>>>>

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
We reconstructed a more comprehensive mutational landscape
of SARS-CoV-2 by mapping deletions onto the ARG
using post-hoc parsimony.
We identified important deletion events
(which are inherited by at least 10,000 samples;
Table~\ref{tab:major_dels}).
All these deletions are known to occur in major VOCs
(https://virological.org/t/preliminary-genomic-characterisation-of-an-emergent-sars-cov-2-lineage-in-the-uk-defined-by-a-novel-set-of-spike-mutations/563;
\cite{Akaishi2023,Li2023indels}).
Eight of these deletions occur in S1-NTD,
a known hotspot for recurrent deletions \cite{McCarthy2021}.
There are three contiguous deletions in NSP6
(4-base at 11288, 5-base at 11283, and 5-base at 11292),
which should have been mapped as two 9-base deletions,
one above the BA.1 origin node and
the other above the BA.2 origin node in the ARG
(Figure~\ref{fig:subgraph_ba1_ba2}),
but they were broken up
during deletion remapping because they had
overlapping four sites at 11288-11292
(which was mapped as the 4-base deletion event).
Furthermore, the placement of these deletion events in the ARG
reflects that of the VOC-associated deletion events
inferred in a previous study \cite{Akaishi2023}.
